COMPUTING ASTROPHYSICAL INFERENCES WITH DIFFUSIVE NESTED SAMPLING

### 2013-11-18 abstract.tex
Abstract done

%%%%%  The title of your contribution in all capital letters.
\title{ASTRONOMICAL INFERENCE WITH DIFFUSIVE NESTED SAMPLING}

@@

Many standard problems in astronomy (e.g. the processing of images to produce
``catalogs'') are best understood as inference problems. In principle at least,
we should calculate the posterior distribution over a suitable hypothesis space,
rather than inventing ad-hoc procedures.

I will describe the Diffusive Nested Sampling method, which is a variant of
Nested Sampling that retains the flexibility of Markov Chain Monte Carlo
exploration of the hypothesis space.

Talk about the StarField project
Talk about Reverberation Mapping
Talk about how Nested Sampling is hard when you need to compress so far
Talk about reversible jump

%%%%%       OPTIONAL: Figure (delete if not used)                                                 
%%%%%       Do not add a figure caption.
%%%%%       The conference program will not be printed in colour so colour figures should not be used.
%%%%%       The example figure is in pdf format.

%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=70mm]{roo}
%\end{center}
%%\caption{blah blah.}
%%\label{fig:roo}
%\end{figure}
%%%%%        End of Figure

%%%%%       OPTIONAL: References (delete  if not used)   
%%%%%       Separate each reference with a blank line.                                                                
\medskip\noindent{References: }

[1] A. Dent et al. Phys. Rev. {\bf E99}, 99998 (2009).

[2] Z. Beeblebrox et al. Phys. Rev. {\bf E99}, 99999 (2009).
%%%%%        END of References

%%%%%       OPTIONAL: Key Words (delete if not used)   
\medskip\noindent{Key Words: }
{Bayesian Inference, Markov Chain Monte Carlo}
%%%%%        END of Key Words

>>>>

%%%%%  The title of your contribution in all capital letters.
\title{COMPUTING ASTROPHYSICAL INFERENCES WITH DIFFUSIVE NESTED SAMPLING}

@@

Many standard problems in astronomy are best understood as inference problems.
In principle at least,
we should calculate the posterior distribution over a suitable hypothesis space,
rather than inventing ad-hoc procedures. I will describe two recent applications
of this idea: i) inferring the properties of active galactic nuclei (AGN)
from reverberation mapping data, and ii) producing ``catalogs'' from images of
crowded stellar fields. The first application allows us to measure the masses
of black holes and to infer the dynamical state of the matter surrounding them,
making the most of the large amount of telescope time required. The second
application has the potential to enable the study of faint objects that are not
even ``detected'' by standard algorithms.
To implement these models we used the ``Diffusive Nested Sampling'' algorithm,
which is a variant of Nested Sampling that was invented at MaxEnt 2009. The
algorithm naturally allows for unknown numbers of parameters since
``reversible jump MCMC'' can be used inside of it. However, there are
computational challenges involved when the posterior distribution is highly
compressed with respect to the prior distribution.

\vspace{1cm}
\medskip\noindent{References: }

[1] {\it Diffusive Nested Sampling}\\
Brendon J. Brewer, Livia B. P{\'a}rtay, Gabor Cs{\'a}nyi,
Statistics and Computing, 2011, 21, 4, 649-656.\\

[2] {\it Geometric and Dynamical Models of Reverberation Mapping Data}\\
Anna Pancoast, Brendon J. Brewer, Tommaso Treu, 2011, ApJ, 730, 139.\\

[3] {\it Probabilistic Catalogs for Crowded Stellar Fields}\\
Brendon J. Brewer, Daniel Foreman-Mackey, David W. Hogg, 2013, AJ, 146, 7.\\
%%%%%        END of References

%%%%%       OPTIONAL: Key Words (delete if not used)   
\medskip\noindent{Key Words: }
{Astrophysics, Bayesian Inference, Markov Chain Monte Carlo}
%%%%%        END of Key Words
